Towards Automated Refactoring of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-07-01 sattose/paper.tex
Magiel

\subsection{Advantages of clone classes over clone pairs}\label{chap:cloneclasses}
Regarding clone detection, there is a lot of variability in literature whether clone pairs or clone classes should be considered for detection. We decided to focus on clone classes, because of the advantages for refactoring. Clone pairs do not provide a general overview of all entities containing the clones, with all their related issues and characteristics~\cite{fontana2012duplicated}. Although clone classes are harder to manage, they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. Another issue that results from grouping clones by pairs: clone reference amount increases according to the binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), which causes a heavy information redundancy~\cite{fontana2012duplicated}.

@@

To limit the negative impact of this tradeoff on the design of the system, we formalized as threshold for the variability between fragments. The formula by which this threshold is calculated is displayed in equation \ref{eq:type2r}. In this formula, \textit{different expressions} refers to the amount of literals, variables and method calls that differ from other clone instances in a clone class. We divide this by the total number of tokens in the clone instance. Based on this threshold, we decide whether a clone should be considered for refactoring.

@@

\begin{itemize}
  \item \textbf{The difference in statements must bridge a gap between two clones that were valid by the original thresholds}. This entails that, different from type 3 clones, the difference in statements cannot be at the beginning or the end of a cloned block. It is rather somewhere within, as it must bridge two existent clones.
  \item \textbf{The size of the gap between two clones is limited by a threshold.} This threshold is calculated by taking the percentage of the amount of statements in the gap over the amount of statements that both clones that are being bridged span. This is displayed in equation \ref{eq:type3r}.
  \item \textbf{The gap may not span a partial block.} To make sure that the T3R clone can be refactored, we do not allow the gap to span a part of a block, for instance the declaration and a part of the body of a for-loop. The reason for this is that it is not possible to wrap a partially spanned block in a single conditional statement. We could however use multiple conditional blocks (one for each block spanned), but due to the detrimental effect on the design of the code (as each conditional block adds a certain complexity), we decided not to allow this for T3R clones.
\end{itemize}

@@

To find such a threshold, we measured the amount of clone classes found for a certain number of tokens. We then looked at the amount of nodes (statements/declarations) that these clones span. The resulting chart is displayed in Fig.~\ref{fig:tokenthreshold}.

\begin{figure}[H]
  \caption{Amount of times a clone with a certain amount of tokens has a certain amount of nodes.}
    \includegraphics[width=1\columnwidth]{img/TokenThreshold}
  \label{fig:tokenthreshold}
\end{figure}

@@

In this figure the amount cloned statements per clone type are displayed. The difference between T1R and T1 is small (10.9\%), because most often textually equal code is also functionally equal. The difference between T2R and T2 is bigger (34.7\%), mainly because we allow no variability in used types. T3R and T3 are similar to T2R and T2, because our dataset does not have so many gapped clones for the thresholds used.

@@

The results are shown in Table~\ref{table:locations}. Our results indicate that, quite significantly, most clones are found at method level. The amount of clones found in interfaces and enumerations is very low.

@@

We measured the amount of clones that can be refactored through method extraction (without additional transformations being required). Our results are displayed in Table \ref{table:refactorability}. In this table we use the following categories:
\begin{itemize}
    \item \textbf{Can be extracted:} This clone is a fragment of code that can directly be extracted to a method. Then, based on the relation between the clone instances, further refactoring techniques can be used to refactor the extracted methods (for instance ``pull up method'' for clones in sibling classes).
    \item \textbf{Complex control flow:} This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements, obstructing the possibility of method extraction.
    \item \textbf{Spans part of a block:} This clone spans a part of a statement.
    \item \textbf{Is not a partial method:} If the clone does not fall in the ``Partial method'' category of Table \ref{table:contents}, the ``extract method'' refactoring technique cannot be applied.
\end{itemize}

>>>>

\subsection{Advantages of clone classes over clone pairs}\label{chap:cloneclasses}
Regarding clone detection, there is a lot of variability in literature whether clone pairs or clone classes should be considered for detection. We decided to focus on clone classes, because of the advantages for refactoring. Clone pairs do not provide a general overview of all entities containing the clones, with all their related issues and characteristics~\cite{fontana2012duplicated}. Although clone classes are harder to manage, they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. Another issue that results from grouping clones by pairs: the amount of clone references increases according to the binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), which causes a heavy information redundancy~\cite{fontana2012duplicated}.

@@

To limit the negative impact of this tradeoff on the design of the system, we formalized as threshold for the variability between fragments. The formula by which this threshold is calculated is displayed in equation \ref{eq:type2r}. In this formula, \textit{different expressions} refers to the number of literals, variables and method calls that differ from other clone instances in a clone class. We divide this by the total number of tokens in the clone instance. Based on this threshold, we decide whether a clone should be considered for refactoring.

@@

\begin{itemize}
  \item \textbf{The difference in statements must bridge a gap between two clones that were valid by the original thresholds}. This entails that, different from type 3 clones, the difference in statements cannot be at the beginning or the end of a cloned block. It is rather somewhere within, as it must bridge two existent clones.
  \item \textbf{The size of the gap between two clones is limited by a threshold.} This threshold is calculated by taking the percentage of the number of statements in the gap over the number of statements that both clones that are being bridged span. This is displayed in equation \ref{eq:type3r}.
  \item \textbf{The gap may not span a partial block.} To make sure that the T3R clone can be refactored, we do not allow the gap to span a part of a block, for instance the declaration and a part of the body of a for-loop. The reason for this is that it is not possible to wrap a partially spanned block in a single conditional statement. We could however use multiple conditional blocks (one for each block spanned), but due to the detrimental effect on the design of the code (as each conditional block adds a certain complexity), we decided not to allow this for T3R clones.
\end{itemize}

@@

To find such a threshold, we measured the number of clone classes found for a certain number of tokens. We then looked at the number of nodes (statements/declarations) that these clones span. The resulting chart is displayed in Fig.~\ref{fig:tokenthreshold}.

\begin{figure}[H]
  \caption{Amount of times a clone with a certain number of tokens has a certain number of nodes.}
    \includegraphics[width=1\columnwidth]{img/TokenThreshold}
  \label{fig:tokenthreshold}
\end{figure}

@@

In this figure the number of cloned statements per clone type are displayed. The difference between T1R and T1 is small (10.9\%), because most often textually equal code is also functionally equal. The difference between T2R and T2 is bigger (34.7\%), mainly because we allow no variability in used types. T3R and T3 are similar to T2R and T2, because our dataset does not have so many gapped clones for the thresholds used.

@@

The results are shown in Table~\ref{table:locations}. Our results indicate that, quite significantly, most clones are found at method level. The number of clones found in interfaces and enumerations is very low.

@@

We measured the number of clones that can be refactored through method extraction (without additional transformations being required). Our results are displayed in Table \ref{table:refactorability}. In this table we use the following categories:
\begin{itemize}
    \item \textbf{Can be extracted:} This clone is a fragment of code that can directly be extracted to a method. Then, based on the relation between the clone instances, further refactoring techniques can be used to refactor the extracted methods (for instance ``pull up method'' for clones in sibling classes).
    \item \textbf{Complex control flow:} This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements, obstructing the possibility of method extraction.
    \item \textbf{Spans part of a block:} This clone spans a part of a statement.
    \item \textbf{Is not a partial method:} If the clone does not fall in the ``Partial method'' category of Table \ref{table:contents}, the ``extract method'' refactoring technique cannot be applied.
\end{itemize}
